---
title: "hw1_zejie8913006"
output: pdf_document
date: "2023-10-11"
---

## 1. Deterministic and stochastic trends. Stationarity.
```{r}
cat("Answer : A\n")
cat("Explanation:\nI is true because deterministic trend is a function estimated
    by regression, and it changes slowly, making it extrapolated.\nII is false 
    because linear models include trends, which violate the rule of a constant 
    mean required for stantionary time series.\nIII is fasle because the 
    autocovariance function can have values greater than 1 in magnitude. ")
```
## 2. Random walk and stationarity 
```{r}
# answer on below
```

## 3. Calculation of sample acf. 
```{r}
# Create a data frame with a stock prices during each day
Xt <- data.frame(
    Number = 1:10,
    price = c(538, 548, 528, 608, 598, 589, 548, 514, 501, 498)
)

# Calculate the mean of the "Value" column
mean_Xt <- mean(Xt$price);mean_Xt

## To estimate autocorrelation at lag 3
lag <- 3
n <- 10

# To compute the covariance of Xt and X(t+3)
autocov_3 <- sum((Xt$price[1:(n - lag)] - mean_Xt) * (Xt$price[(lag + 1):n] - mean_Xt)) ;autocov_3

# To compute the variance of Xt
variance <- sum((Xt$price - mean_Xt)^2) ;variance

# Computing autocorrelation by dividing covariance with variance
autocorr_3 <- autocov_3 / variance
cat("autocorr_3 :", autocorr_3, "\n")
# answer on below
```
## 4. Polyroot command in R.
```{r}
root_of_f <- polyroot(c(1, -5))
cat("Root of f(z): ")
cat(root_of_f, sep=", ")

root_of_g <- polyroot(c(1, -0.75, 0.125))
cat("\nRoot of g(z): ")
cat(root_of_g, sep=", ")
# answer on below
```
## 5. Model identification. 
```{r}
# answer on below
```

## 6. Please indicated True/False and explain briefly why.
(a) False
Explanation: for (i), it is not necessary for white noise to have 0 mean as long as it is constant, (ii) is true, (iii) also have be satisfied for meeting the  uncorrelated property. 
(b) True
additional proof on below
## 7. Gaussian White Noise and its square.
(a)
```{r}
set.seed(1)
# Generate 350 observations of the Gaussian white noise Z
Z <- rnorm(350)
plot.ts(Z, xlab = "time", ylab = "Z") 
acf(Z, main = "ACF")
```
(b)
```{r}
# Defining Y using Z
Y <- Z^2
plot.ts(Y, xlab = "time", ylab = "Y") 
acf(Y, main = "ACF")
```
(c)
- Series Z holds constant mean and variance, and series Y also do not shows trend and various variance over time. The value of mean in Z and Y on the graph looks different. Comparing only series graph is not enough for knowing whether it is stationary or not. Additional test of autocorrelation is need.
- There is not noticeable difference in the plots of acf function between Z and Y. Combined with nearly constant mean and variance from series graph, it is highly possible that both Z and Y are stationary since both acf graph exhibits autocorrelations close to zero for all lags except at lag 0, where it is always 1. Y is described as non-Gaussian white noise due to its non-zero mean and white noise properties.
(d)
```{r}
maen_y <-mean(Y);maen_y
std_Y <- sd(Y);std_Y
```
The calculations indeed support my observations in (c), by considering non-zero but constant mean and variance, and uncorrelated property. More proof on below.  
```{r}

```

