---
title: "lab1_zejie"
output: pdf_document
date: "2023-10-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. read in the data
```{r}
births <- scan("http://robjhyndman.com/tsdldata/data/nybirths.dat")
```
# 2. Convert it into time series format
```{r}
births_ts = ts(births, start = c(1946,1), frequency = 12)
births_ts
```
# 3. Plot the data versus time
```{r}
plot(births_ts, main = "Time Series from Year 1946 to 1969")
```
## 4. Plot the aggregated data versus time
```{r}
library(xts)
ts.yearly <- apply.yearly(as.xts(births_ts), FUN = mean)
ts.yearly
```
# Plot the Aggregated yearly data:
```{r}
plot(ts.yearly, main = "Yearly Time Series from Year 1946 to 1959")
```
## 5 Plot the data versus time index number from 1, 2, ...n. Add regression line and mean line to the plot and print sample size n.
```{r}
plot(1:length(births),births, main =
   "Time Series from Year 1946 to 1959", type = 'l',xlab='index')
index = 1: length(births)
trend <- lm(births ~ index)
abline(trend, col = "red")
abline(h = mean(births), col = "blue")
```
```{r}
length(births)
```

## 6. State whether the data looks stationary. 
```{r}
birthstimeseriescomponents <- decompose(births_ts)
plot(birthstimeseriescomponents)
```
```{r}
wn = rnorm(1000,0,1)
plot(1:length(wn),wn, main =
   "White Noise Plot", type = 'l',xlab='index')
trend <- lm(wn ~ time(wn))
abline(trend, col="red")
abline(h=mean(wn) , col='blue')
```
## 7. Does 1946-1948 data have the same pattern as the rest of the data? If no, please cut the start of the data e.g. your data will start at Jan, 1948.
```{r}
select <- c(1:24)
#eliminate data from year Jan 1946 to Dec 1947 and we therefore now have 12 years(144 obse 
birth_split <- births[-c(select)]
#use the first 11 years(132 observations) for training
training = birth_split[1:132]
#use the last 1 year(12 observations) for testing
testing = birth_split[133:144]

```
## 8. Generating Uniform distribution and calculate the mean.
```{r}
observations = runif(1000,-1,1)
mean(observations)
```

